---
title: 'ML #4'
output: html_document
---

```{r}
library(ISLR)
library(tidyverse)
library(factoextra)
library(ggplot2)
library(readr)
library(corrplot)
library(esquisse)
library(caret)

Universities <- read_csv("Universities.csv")
View(Universities)
head(Universities)
summary(Universities)

Universities$Acceptance_rate <- Universities$`# appl. accepted`  /Universities$`# appli. rec'd`*100   #added column for acceptance rate (#Application Accepted /# Application received)

Universities[duplicated(Universities$`College Name`),]   #verify no duplicated rows

colMeans(is.na(Universities))   #Calculate missing values per column

Universities_complete <- Universities[complete.cases(Universities), ] #Remove all records with missing measurements from the dataset
colMeans(is.na(Universities))  

```


Normalizing Continuous Variables
```{r}
Numeric_subset <- Universities_complete[, c(4:21)]    #Making a susbet for numeric variables for normalization 

corrplot(cor(Numeric_subset) , method= "color")  #Checking for correlation relations between variables

Universities_normalized <- scale(Numeric_subset)  #Normalizing dataset using Z-score

Distance <- get_dist(Universities_normalized) 

fviz_dist(Distance)  #visualization for distance between universities using normalized data as distance measure is sensitive to scale


set.seed(123)
Cluster_k5 <- kmeans(Universities_normalized, centers= 5, nstart =  20)  #Using k-means to cluster our data into 5 clusters with number of 20 restarts(k=5)

Cluster_k5$size  # shows the number of observations(Universities) in each of the  5 clusters 
Cluster_k5$cluster     #show cluster membership


Cluster_k5$centers  #shows the centroid for each cluster. we can see here that cluster #4 has the highest number of application received and accepted where Cluster #1 has the lowest number of applications 
Cluster_k5$withinss    #shows the within-cluster sum of squares. we can see here that Cluster #1 has the highest WSS, highest dispersion .


fviz_cluster(Cluster_k5, data =  Universities_normalized)  #Visualizing the 5 clusters, we can see how some clusters are overlapping. Next we find optimal K.

#Now we Use Elbow Method to find the optimal number of clusters 
fviz_nbclust(Universities_normalized, kmeans, method =  "wss")  #looks like moving from 2 to 3 clusters tighten the Total WSS

#Using Silhouette Method confirms that K=3 is the optimal number of clusters.
set.seed(125)
fviz_nbclust(Universities_normalized, kmeans, method =  "silhouette")

```

Now we re-run the model using the optimal number of clusters k=3 and compare clustering results and describe each cluster 
```{r}

set.seed(124)
Cluster_k3 <- kmeans(Universities_normalized, centers=3 , nstart = 25)
Cluster_k3

Cluster_k3$size   #number of universities in each cluster (Cluster #1 has the highest numbers of Universities)

Cluster_k3$withinss   #Cluster number 3 has the lowest within_cluster sum of squares and cluster #1 has the highest WSS

#Cluster centroids shows that cluster #3 has the highest number of applications recieved and accepted , highest number of enrollment for new students, lowest in-state and out-of-state tuition, second highest percentage rate of new students from top 10% and top 25%, and second highest graduation and acceptance rate. on the other hand Cluster #1 has the lowest number of application received and accepted,  lowest % of new students enrollment,lowest percentage from top 10% and top 25%, second highest in-state-and out-of-state tuition, lowest graduation rate, and highest acceptance rate.   Cluster #2 has the highest in-state and out-of-state tuition, lowest acceptance rate, highest graduation rate, highest percentage of students from top 10% and top 25%
Cluster_k3$centers #shows cluster centroids

Cluster_k3$cluster   #show cluster membership for each University(Observation)

as.matrix(dist(Cluster_k3$centers))  #viewing distance between clusters. It shows that clusters 1,2 are the closest 

fviz_cluster(Cluster_k3, data =  Universities_normalized)  #Visualizing the 3 clusters 


#Now we run summary statistics on original data

#we look first at the average for each variable by cluster. Cluster 3 has the highest average number of application and new enrollments with second highest acceptance rate and graduation rate. It also has the lowest average in-state and out-of-state tuition. 
aggregate(Numeric_subset, by= list(Cluster_k3$cluster), FUN= mean) 

#we  also look at median, min, max and standard deviation for each cluster
Summary_Uni<- Numeric_subset %>%  
  mutate (Clusters = Cluster_k3$cluster) %>% 
  group_by(Clusters) %>% 
  summarise_all(list( median= "median",min= "min",max= "max",sd= "sd"))

Summary_Uni


#Cluster 2 has the highest median in-state-Tuition followed by cluster 2 
ggplot(Summary_Uni) +
 aes(x = Clusters, weight = `in-state tuition_median`) +
 geom_bar(fill = "#0c4c8a") +
 theme_minimal()

#Cluster 3 has the highest number of application received followed by cluster 2
ggplot(Summary_Uni) +
 aes(x = Clusters, weight = `# appli. rec'd_median`) +
 geom_bar(fill = "#0c4c8a") +
 theme_minimal()

#cluster 2 has the highest number of students from top 10%
ggplot(Summary_Uni) +
 aes(x = Clusters, weight = `% new stud. from top 10%_median`) +
 geom_bar(fill = "#0c4c8a") +
 theme_minimal()

```


Now we use categorical variables to characterize clusters 
```{r} 
Cluster_Categorical <- cbind( Universities_complete, Clusterss= Cluster_k3$cluster)

#We have a total number of 343 Private universities in dataset and 128 private schools. We need to consider that we have more private than public schools. Public schools only account for around 27%.
nrow(filter(Universities_complete,  Universities_complete$`Public (1)/ Private (2)`==1))
nrow(filter(Universities_complete,  Universities_complete$`Public (1)/ Private (2)`==2))

#Visualize clusters size. We can see cluster 3 has the biggest number of universities
ggplot(Cluster_Categorical) + aes(x = Clusterss) + geom_bar(fill = "#2171b5") +
 labs(x = "Cluster Category", y = "Count") +
 coord_flip() + theme_light()
 

#Overall Public universities have lower distribution of In_State Tuition across all clusters than private Universities
ggplot(Cluster_Categorical) +
 aes(x =as.factor (Clusterss), y = `in-state tuition`, fill = as.factor(`Public (1)/ Private (2)`) )+
 geom_boxplot() + scale_fill_brewer(palette = "BrBG") +
 labs(x = "Clusters", y = "In_State Tuition", title = "Distribution of In_State Tuition Per Cluster for public and private Universities") + theme_light()


#In this graph we see that Universities in Cluster 1 have higher acceptance rate than other 2 clusters.
ggplot(Cluster_Categorical) +
 aes(x = Acceptance_rate, fill =as.factor (Clusterss)) +
 geom_histogram(bins = 27L, position = "dodge") +
 labs(title = "Acceptance Rate Per Cluster")


#Let's look at the number of public Vs Private Universities Per cluster. Cluster 2 is dominated by Private schools, whereas cluster 3 is dominated by Public schools . Cluster 2 has mixed numbers of both public and private

ggplot(Cluster_Categorical) +
 aes(x =as.factor (Clusterss), fill = as.factor(`Public (1)/ Private (2)`)) +
 geom_bar() + scale_fill_brewer(palette = "BuPu") +
 labs(x = "Clusters", y = "Number of Universities ", title = "Number of Public Vs Private Universities Per Cluster") +
 coord_flip() + theme_minimal()


#Students in public schools tend to spend more on personal expenses than students in private schools
ggplot(Cluster_Categorical) +
 aes(x =as.factor (`Public (1)/ Private (2)`), y = `estim. personal $`) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal()


#However Private schools in cluster 1 have higher acceptance rates than private schools in other cluster. Acceptance rates for public schools in cluster 2 are lower than res of the clusters
ggplot(Cluster_Categorical) +
 aes(x =as.factor (`Public (1)/ Private (2)`), y = Acceptance_rate, fill =as.factor (Clusterss)) +
 geom_boxplot() + scale_fill_hue() + theme_minimal()


#Added fees for public universities are higher that private ones. 
ggplot(Cluster_Categorical) +
 aes(x =as.factor (`Public (1)/ Private (2)`), y = `add. fees`, fill = as.factor(Clusterss)) +
 geom_boxplot() + scale_fill_hue() + theme_minimal()

#External information that would explain the clustering is the fact that public tend to have lower tuition fees and higher applications with medium-levels in graduations rates. Whereas Private schools tend to be more expensive and have higher graduation rates with low acceptance rate. Also, the income of the students in both private and public schools could add more insight into the reason why they seem to spend more than students in private schools. If we have more information about the quality of education we can look into why students from top 10% and top 25% are more likely to be in private schools than in public.

```


Consider Tufts University , compute Euclidean distance for this record from each of the clusters . which clusters is it close to? and impute missing values by taking the average. 
```{r}
 #this row only have one missing value in #PT undergrad column
Tufts_Uni <- filter(Universities, Universities$`College Name` == "Tufts University")
as.matrix(Tufts_Uni)

#First we normalize the data
Uni_df <- Universities
Norm_Uni_df <- preProcess(Uni_df[, -c(1:3)], method = c  ("scale","center"))
Tufts <- filter(Uni_df, Uni_df$`College Name` == "Tufts University")
Tufts[, -c(1:3)] <- predict(Norm_Uni_df, Tufts[, -c(1:3)])
Tufts

#Now we find the distance between this record and all 3 clusters centroids.
dist(rbind(Tufts[,-c(1:3)],Cluster_k3$centers[1,]))
dist(rbind(Tufts[,-c(1:3)],Cluster_k3$centers[2,]))  #Distance = 3.42. Tufts University is closest to cluster 2.
dist(rbind(Tufts[,-c(1:3)],Cluster_k3$centers[3,]))

#Impute the missing value in Tufts university's record by taking the mean of the cluster
Cluster_2 <- Cluster_Categorical %>% 
  filter(Clusterss == 2)

mean(Cluster_2$`# PT undergrad`)  #the imputed missing value is 297.49

```













